---
title: "WUP_04"
author: "Tyler Hamblin"
date: "Fall 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
mydata <- read.csv(file = "GSScsv.csv")
```

```{r eval=FALSE}
mydata <- mydata %>%
mutate(across(where(is.character), as.factor))
```



# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

With the help of ChatGPT and the helper files under the CandP zip we will be able to decide whether or not the WORDSUM test scores are affected by the level of education of the test taker. 



# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r include=TRUE,echo=FALSE}
ggplot(data = mydata, aes(x = DEGREE, y = WORDSUM) ) +
geom_boxplot ( notch=FALSE , outlier.colour = "red" , outlier.size = 2 , fill = "darkgray" ) +
stat_boxplot ( geom = 'errorbar' ) +
labs ( x = "DEGREE" , y = "WORDSUM" , title = "WORDSUM Explained By DEGREE" ) +
theme ( plot.title = element_text (hjust = 0.5, size = 16) )
```

The first quartile of both the High School and Lt High School are still below the medians of all other groups suggesting that the level of education does affect scores. This is not enough evidence however and more tests need to be run to confirm this initial theory. 

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r include=TRUE,echo=FALSE}
favstats( WORDSUM ~ DEGREE , data= mydata )
```

Under the LT High School and High School categories the mean and median are significantly lower than the means and medians of the university level data points. This continues to confirm the results from the box plot as students from the less advanced levels of education seem to be underperforming when compared to the rest. 


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r include=TRUE,echo=FALSE}
model1 <- aov( WORDSUM ~ DEGREE , data =  mydata )
conf_interval <- confint(model1, level= 0.98)
conf_interval
summary.aov(model1)
```

```{r include=TRUE,echo=FALSE}
par(mar = c(3, 15, 2, 2))
plot(TukeyHSD(model1,conf.level = 0.99),las=2)
```

As seen in the one way anova, the confidence interval test, and the graphical version of the family-wise confidence level analysis, the LT High School and the High School seem to be the lowest performers on the WORDSUM test. The categories containing either High School or LT High School place their median the farthest away from the 0 mark suggesting a significance in statistical data. This confirms that as a students education continues they theoretically will begin to achieve higher scores on this exam. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->




